\documentclass[class=scrartcl, crop=false]{standalone}

\usepackage[sexy]{evan}
\usepackage{cole}


\date{2020-03-10}


\begin{document}

\section{Lecture 03-10}

\begin{example}
  $f_n(x) = x^n, x \in [0, 1]$. For each $n$ fixed. $f_n$ is continuous.
  \begin{gather*}
    \lim_{n \to \infty}f_n(x)
    \begin{cases}
      0, & 0 \leq x < 1 \\
      1, & x = 1
    \end{cases} 
  \end{gather*} 
\end{example} 

Question. When does $\lim_{n \to \infty}f_n(x)$ exist and is continuous.

We need to impose some further conditions to ensure that $\lim f_n(x)$ is continuous.
\\\\
Suppose $f(x) = \lim_{n \to \infty}f_n(x)$. $f$ is continuous at $x_0$ iff
\begin{gather*}
  \lim_{x \to x_0}f(x) = f(x_0) = \lim_{n \to \infty}\lim_{x \to x_0}f_n(x) \\
  = \lim_{x \to x_0}\lim_{n \to \infty} f_n(x)
\end{gather*} 
So interchanging the two limits and it is still equal.
\\\\
\begin{definition}[Uniform Convergence]
  Let $\{f_n\}$ be a sequence of functions on $X$. We say that $\{f_n\}$ is uniformly convergent in $X$ to $f$ if $\forall \epsilon > 0, \ \exists N$ such that $|f_n(x) - f(x)| < \epsilon, \ \forall n \geq N$.
  \begin{note}
    The choice of $N$ is independent of $x \in X$. It depends only on $\epsilon > 0$.
  \end{note} 
\end{definition} 

\begin{example}
  Let's revisit the example $\{x^n\}$. 
  \begin{gather*}
    f_n \to f \ \text{uniformly continuous} \ \Rightarrow f_n(x) \to f(x) \ \text{pointwise} \ 
  \end{gather*} 
  \\\\
  $|f_n(x) - f(x)| < \epsilon$ if $n \geq N(\epsilon)$.
  \begin{enumerate}
    \ii
    When $x = 1$, this is true $\forall n$.
    \ii
    The problem is $0 \leq x < 1$.
    \begin{gather*}
      \Rightarrow |x^n| < \epsilon, \ \forall n \geq N(\epsilon)
    \end{gather*} 
    For all $N$ fixed, we can always find $x_n$, such that $x_n^n \to 1, \ n \to \infty$.
  \end{enumerate}
\end{example} 

\begin{lemma}
  $f_n \to f$ is uniformly continuous iff $\forall \epsilon > 0, \ \exists N$ such that $\forall m > n \geq N$, $|f_m(x) - f_n(x)| < \epsilon, \forall x \in X$.
  \begin{proof}
    $\Rightarrow$. Suppose $f_n \to f$ is uniformly continuous. i.e.
    \begin{gather*}
      \exists N, \ \text{such that} \ \forall n \geq N, \\
      |f_n(x) - f(x)| < \frac{\epsilon}{2}, \ \forall x \in X
    \end{gather*} 
    Then
    \begin{gather*}
      m > n \geq N \\
      |f_m(x) - f_n(x)| \leq |f_m(x) - f(x)| + |f(x) - f_n(x)| \leq \epsilon / 2 + \epsilon / 2 = \epsilon
    \end{gather*} 
    $\Leftarrow$. If $(*)$ holds, then $\{f_n(x)\}$ is cauchy. $f_n(x) \to f(x), \ \forall x \in X$ pointwise because $\RR$ is complete.
    \begin{gather*}
      |f_n(x) - f(x)| \leq |f_n(x) - f_m(x)| + |f_m(x) - f(x)| \leq \epsilon + |f_m(x) - f(x)|
    \end{gather*} 
    Let $m \to \infty$
     \begin{gather*}
       \epsilon + |f_m(x) - f(x)| \leq \epsilon
    \end{gather*} 
  \end{proof} 
\end{lemma} 

\begin{theorem}
  Suppose $\{f_n\}$ is a sequence of continuous functions defined in $X$. Suppose $f_n \to f$ is uniformly convergent in $X$. Then $f$ is continuous in $X$.
\end{theorem} 

\begin{theorem}
  Suppose $\{f_n\}$ is a sequence of continuous functions uniformly convergent to $f$ in $X$, $x_0 \in X$ is a limit point of $X$.
  Suppose then
  \begin{gather*}
    \lim_{x \to x_0} f_n(x) = A_n
  \end{gather*}
  \begin{gather*}
    \lim_{x \to x_0}f(x) = \lim_{n \to \infty}A_n = A
  \end{gather*}
  \begin{proof}
    By cauchy $(*)$, $\forall \epsilon > 0, \exists N, \ |f_m(x) - f_n(x)| \leq \epsilon, \ \forall m > n \geq N, \forall x \in X$.
  \end{proof} 
  Let $x \to x_0 \Rightarrow |A_m - A_n| < \epsilon, \ m, \geq N \Rightarrow \{A_n\}$ is cauchy. $A_n \to A$.
  \begin{gather*}
    |f(x) - A| \leq |f(x) - f_n(x)| + |f_n(x) - A_n| + |A_n - A|
  \end{gather*} 
  We need to find $\delta > 0$ such that $\forall x \in B_\delta(x_0), \ |f(x) - A| < 3\epsilon$.
  \\
  Set $n = N \Rightarrow |f(x) - A| \leq 2\epsilon + |f_N(x) - A_N|$. By the assumption, $\lim_{x \to x_0} f_N(x) = A_n \Rightarrow \exists \delta > 0$ such that $|f_N(x) - A_n| < \epsilon, \ \forall x \in B_\delta (x_0)$. This is the $\delta$ we are looking for.
\end{theorem} 

\begin{theorem}[M-test]
  When $f_n \to f$ is uniformly continuous in $X$.
  \\\\
  $|f_n(x)| \leq M_n, \ \forall x \in X$, consider $\sum_{n = 1}^\infty f_n(x)$ is uniform if $\sum_{n = 1}\infty M_n$ is convergent.
\end{theorem} 
\end{document}
