\documentclass[class=scrartcl, crop=false]{standalone}

\usepackage[sexy]{evan}
\usepackage{cole}


\date{2020-03-12}


\begin{document}

\section{Lecture 03-12}

\subsection{Amortized Time}

\begin{definition}[Amortized Time]
  Amortized Time $\coloneqq$ Actual Time $+ \Delta \phi_{\text{Potential}}$ 
  \begin{gather*}
    \phi \geq 0
  \end{gather*} 
\end{definition} 

\begin{note}
  If $\phi(\text{before}) = 0$, then Amortized time $\geq$ Actual time.
\end{note} 

\begin{example}[Lazy Delete]
  \begin{itemize}
    \ii
    Fake / marked items.
    \ii
    At 50\% "fake", reconstruct the data structure.
  \end{itemize} 
  Let $\phi = 2 \cdot$ \# Fake Elements. Then:
  \begin{center}
  \begin{table}[]
  \begin{tabular}{|l|l|l|l|}
  \hline
              & Actual & \textbackslash{}Delta \textbackslash{}phi & AM    \\ \hline
  Insert      & logn   & 0                                         & log n \\ \hline
  Lazy Delete & 1      & 2                                         & 3     \\ \hline
  Reconstruct & n      & -n                                        & 0     \\ \hline
  \end{tabular}
  \end{table}
  \end{center} 
  Actual time: \ul{Cost} of $t$ operations (Insert, Delete) starting from an empty tree % review this
  \begin{gather*}
    \leq t \log t + 3t
  \end{gather*} 
\end{example} 

\begin{example}[Counting to $n$ in the bit model]
  Let $\phi$ be the number of ones.
  \\\\
  Actual time take  $k + 1$ where $k$ is the number of ones before the first zero reading from left to right.
  \\
  $\Delta \phi = -k + 1$. \\
  So Amortized time $= 2$ for each increment.
  \\\\
  Therefore total amortized time is $= 2n$, so actual time is $\leq 2n$.
\end{example} 

\subsection{Fibonacci Heap}

Operations. % Clean up table
  \begin{table}[]
  \begin{tabular}{|l|l|l|l|}
  \hline
              & Binary Heap & Fibonacci Heap (Amortized) & AM    \\ \hline
    Insert      & logn   & 1 & log n \\ \hline
    Delete & logn      & \log n& 3     \\ \hline
    Delete Min & logn      & \log n& 0     \\ \hline
    Decrease Key & logn      & 1 & 0     \\ \hline
    Meld (Join) & logn * logm      & 1& 0     \\ \hline
  \end{tabular}
  \end{table}

General Stucture is a bunch of double linked lists connected to another another, along with a "root list" and pointer to the min in the "root list".
\\\\
\begin{enumerate}
  \ii
  Insert(x, F). Add another "subtree" to the root list.
  \ii
  Meld($F_1, F_2$). Link the root lists of the two trees together.
  \ii
  Decrease key ($x, k, F$ ). Insert $x$ tree into root list and change the key.
  \ii
  Delete Min. First return min and remote it from the root list. Then move all children of the min into the root list. Then pick a new min.
  \ii
  Delete. First identify the element in log n time, then move all the children to the root list.
  \ii
  Clean up (Root list).
  \begin{enumerate}
    \ii
    First empty the root list into a stack $S$.
    \ii
    Then while $|S| > 0$, 
  \end{enumerate} 
\end{enumerate} 
We will ensure that max degree $= O(\log n)$. Proof later.
Let $\phi = \alpha $ size of root list + $\beta $ number of marked items.


\end{document}
