\documentclass[../598comp.tex]{subfiles}

\graphicspath{ {./lectures/images/}{./images/} }

\date{05-27}

\begin{document}

\section{05-27}

\subsection{Universal Functions}

\begin{definition}
  A binary function $U: \NN^2 \to \NN$ is said to be \ul{universal} for the
  class of \ul{computable} unary functions if
  \begin{enumerate}
  \item 
    $\forall n, \ U_n: x \mapsto U(n, x)$ is computable. $U_n$ is called a
    \ul{section} of $U$. This is called currying, when you split a function that
    takes multiple parameters into nested unary functions.
  \item
    $\forall$ unary computable $f: \NN \to \NN, \ \exists n$ such that $U_n =
    f$, i.e. $\forall x \ U_n(x) = f(x)$
  \end{enumerate}
  Note that in this definition $U$ doesn't have to be computable.
\end{definition}

\begin{note}
  The list of programs is countable.
\end{note}

\begin{theorem}
  There is a binary \ul{computable} function $U: \NN \times \NN \to N$ such that
  $U$ is a universal function for all unary computable functions. i.e. one
  turing machine that can simulate all the others.
  \begin{proof}
    Consider your favorite programming language (YFPL), and enumerate all legal
    programs, $p_1, p_2, \dots, p_n, \dots$

    $U(n, x) = p_n(x)$. So $U$ is an ``interpreter'' while $n$ is the code of
    the algorithm.
  \end{proof}
  \begin{note}
    ``Code'' comes from the days of early computability theory because every
    program could be coded up as a number.
  \end{note}
\end{theorem}

\subsection{Total Computable Universal Function}

Does there exist a \ul{total} computable universal function for the class of
\ul{total} computable unary functions. Total means it will be defined on all
inputs, so everything must terminate. No!

Let $U$ be any total computable function of two arguments. Define $d(n) = U(n,
n) + 1$. $\forall n, \ d(n) \neq U_n(n)$ so $\forall d \neq U_n$. Can't
guarantee all terminating algorithms, or must allow some options to not terminate.

Think about why doesn't this argument work for partial functions? Because $U(n,
n)$ might be undefined, in which case $U(n, n) + 1$ is still undefined.

\subsection{Compositional Programming}

We want to program \ul{Compositionally}. If $f, g$ rae computable functions,
then $g \circ f$ is also computable. The map that figures out $g \circ f$ should
be total computable.

\begin{definition}
  Let $S$ be any countable set. A map $\nu: \NN \to S$ is called a
  \ul{numbering} of $S$ if $\nu$ is surjective.

  A value of $n$ such that $\nu(n) = s$ is called a code number for $s$.

  Note the indefine article. There can be multiple such $n$ for a given element.

  We want to show that for the ``right kind'' of universal function $U$, there
  is a computable function with the following properties.
  \begin{gather*}
    c: \NN \times \NN \to \NN \\
    \forall p, q, x \in \NN, \ (U_p \circ U_q)(x) = U(p, U(q, x)) = U(c(p, q),
    x) = U_{c(p, q)}(x)
  \end{gather*}
\end{definition}

\begin{note}
  A set $S$ is computable if there is a \ul{total} computable function $f$, such
  that $f(n) = 1$ if $n \in S$ and $f(n) = 0$ if $n \notin S$.

  Note that computable function doesn't have to be ``computable'' set. Only total
  computable function.
\end{note}

\begin{definition}
  Let $U$ be a universal computable function. It is called a \ul{Godel}
  universal function if $\forall$ binary computable functions $V$, $\exists$ a
  total computable unary function $\sigma: \NN \to \NN$ such that $\forall m, x
  \in \NN, \ V(m, x) = U(\sigma(m), x)$ ($\sigma$ will depend on $V$).
\end{definition}

I stopped taking notes because I realized there was a handout online with
detailed notes on today's lecture.

\subsection{Primitive Recursive Functions}

Godel. These roughly correspond to a programming language with \ul{bounded} search. i.e
can't use while loops, only loops that run a set number of times.

Fortran for loops for example. Provably terminating.

Ackerman gave an example of a provably terminating function that was not
primitive recursive.

This makes sense based on the proof above, whereby it's impossible to produce a
universal total computable function for the class of total computable unary functions.

Kleene: PRF + unbounded search gives partial recursive functions. They include
the power of while loops. Proved that these are equivalent to turing machines
and lambda calculus.

\subsection{Degrees of Unsolvability}

Online handout.

\end{document}
